A Closer Look at Network Flow on Evolving Graphs

### 2018-05-16 paper.tex
update temporal closeness and pagerank part

% closeness centrality
Let $d_{ij}$ be the length of a temporal shortest walk from $i$ and $j$ in an evolving graph. Then
the closeness centrality of a node $i$ is defined
$$
C_i^{closeness} = \frac{N-1}{\sum_j d_{ij}}.
$$
Note that this depends on how we define the length of temporal shortest walk. The closeness centrality can takes account of only static edges, only causal edges, or both.
The definition of temporal closeness centrality is similar to temporal katz. The only difference is that here we traverse the shortest paths not just walks on evolving graphs.

\begin{lstlisting}
function temporal_closeness(g::AbstractEvolvingGraph, start::TimeNode)
  v = start
  level = Dict(v => 0)
  i = 1
  fronter = [v]
  while length(fronter) > 0
      next = []
      for u in fronter
          for v in forward_neighbors(g, u)
              if !(v in keys(level))
                  td = temporal_distance(start, u)
                  level[v] = i + td
                  push!(next, v)
              end
          end
      end
      fronter = next
      i += 1
  end

  total_scores = sum(values(level))
  return total_scores > 0. ? (length(level) - 1)/total_scores : 0.
end
\end{lstlisting}
For Figure \ref{fig:katz_eg2} the scores are
\begin{lstlisting}
Dict{Any,Any} with 7 entries:
TimeNode(B, 3) => 0.0
TimeNode(B, 1) => 0.27027
TimeNode(A, 3) => 1.0
TimeNode(B, 2) => 1.0
TimeNode(C, 1) => 0.259259
TimeNode(A, 2) => 0.538462
TimeNode(C, 2) => 0.416667
\end{lstlisting}
For Figure \ref{fig:katz_eg3}, the scores are
\begin{lstlisting}
Dict{Any,Any} with 7 entries:
TimeNode(B, 100) => 0.0
TimeNode(A, 2)   => 0.0636364
TimeNode(A, 100) => 1.0
TimeNode(C, 2)   => 0.0458716
TimeNode(B, 1)   => 0.0746269
TimeNode(C, 1)   => 0.0564516
TimeNode(B, 2)   => 1.0
\end{lstlisting}

The analysis for temporal communicability betweenness centrality is similar.

@@

We derive a generalization of PageRank on evolving graphs.
The first idea is based on reverse temporal flows and block adjacency matrix.

\subsubsection{Reverse Temporal Flows}

We can not travel back in time. But reversing the direction of temporal flows can help
identify the ``hub'' nodes or the source of the flows. The authority score of a page is
proportional to its importance, and the hub score describes the quality of a page as a link collection of important related pages \cite{kleinberg99}.
Indeed, reverse PageRank has
been studied by \cite{bar08, fogaras03, gleich15}. In reverse PageRank we compute PageRank on the graph with reversed direction, i.e., reverse the direction of each edge $(i,j)$ to $(j, i)$.
Fogaras \cite{fogaras03} shows that Reversed Page Rank scores express hub quality.

Recall a time-preserving walk is a time-ordered sequence of active nodes. Here we define a \emph{reversed temporal walk} to be a reversed time-ordered sequence of active nodes, i.e.,
$t_1 \ge t_2 \ge \cdots \ge t_m$.

@@

We can not simply use static graphs to replace evolving graphs for two main reasons.
First, a path has to preserve the dimension, i.e., certain path is incorrect in the aggregated static graph.
Second, two path between $A$ and $B$ are considered to be one path in the aggregated static graph.

In \cite{chen16}, we derive a block adjacency matrix representation of an evolving graph.
If $E'$ is the set of causal edges and $\hat E$ is the set of static edges then
an evolving graph can be represented as
$$
\bm M_n =
\begin{pmatrix}
A^{[t_1]} & M^{[t_1, t_2]} & \ldots & M^{[t_1, t_n]} \\
0         & A^{[t_2]} & \ldots & M^{[t_2, t_n]} \\
          & \ldots    &        &     \\
0         & 0         & \ldots & A^{[t_n]}
\end{pmatrix},
$$
where $M^{[t_i, t_j]}$ is the matrix whose rows are labeled by $V^{[t_i]}$ and columns are labeled by $V^{[t_j]}$, and whose entries are
$$
  M_{uv}^{[t_i, t_j]} =
  \begin{cases}
    1 & \mbox{if} (u, v) \in E' \\
    0 & \mbox{otherwise}.
  \end{cases}
$$
The adjacency matrix blocks $A^{[t]}$ encode the static edge set $\hat E$, whereas the off-diagonal blocks $M^{[t_i, t_j]}$ encode the causal edge set $E'$.

(Need to justify the benefit of this approach.)
We can apply static graph centrality algorithms to $M_n$ and then take the average of ratings at each each time stamp. For example, we can construct the Google matrix
$G_n$ from $M_n$ and the ranking of the nodes can be found by solving
$$
 r= G_n r.
$$

For Figue \ref{fig:eg_shortest_walk}, we have

\begin{align*}
V & = \{(1,\!t_1), (2,\!t_1), (1,\!t_2), (3,\!t_2), (2,\!t_3), (3,\!t_3)\},\\
\tilde E &= \{((1,\!t_1), (2,\!t_1)), ((1,\!t_2), (3,\!t_2)), ((2,\!t_3), (3,\!t_3))\},\\
E'  &= \{((1,\!t_1), (1,\!t_2)), ((2,\!t_2), (2,\!t_3)), ((3,\!t_2), (3,\!t_3))\}.
\end{align*}

@@

\[
\bm M_3 = \begin{pmatrix}
0 & 1 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 0
\end{pmatrix}
\]

@@

We derive the block google matrix as a combination of block matrix $A$ and its transpose $A^T$.

The block google matrix is defined as

\begin{lstlisting}
function block_google_matrix(g::IntAdjacencyList; alpha = 0.85, balance = 0.75)
  A = full(block_adjacency_matrix(g))
  A = balance * A + (1-balance) * A'
  N, N = size(A)
  p = ones(N)/N
  dangling_weights = p
  dangling_nodes = find(x->x==0, sum(A,2))

  for node in dangling_nodes
      A[node,:] = dangling_weights
  end
  A ./= sum(A,2)
  return alpha * A .+ (1- alpha) * p
end
\end{lstlisting}

We find the ranking for Figure \ref{}

\begin{lstlisting}
9-element Array{Complex{Float64},1}:
 0.12278
0.254264
0.0533627
 0.36275
0.350713
0.406792
0.0533627
0.524112
0.468848
\end{lstlisting}

One potential problem is the computational cost. Since the dimension of the block matrix
is $|V| \times |T|$, where $|V|$ represents the number of nodes and $|T|$ represents the number of time stamps. The block matrix may become very large for an evolving graph with many time stamps.

\section{Experiments}
\label{sec:experiments}

We test the influence of time on evolvign graph centralities.

>>>>

% closeness centrality

Let $d_{ij}$ be the length of the shortest path from $i$ and $j$ in a static graph. Then
the closeness centrality of a node $i$ is defined
$$
C_i^{closeness} = \frac{N-1}{\sum_j d_{ij}},
$$
where $N-1$ is the total number of reachable nodes.
For evolving graphs, we need to consider the shortest temporal path from $(i,t_i)$ to $(j,t_j)$, for any time stamp $t_i \leq t_j$.
Depends on how define the length of shortest temporal path, the closeness centrality can counts static edges, causal edges, or both.
The key difference between Katz centrality and closeness centrality is the fact here we only consider shortest temporal paths, not just every temporal walks.
We could apply BFS to find the shortest temporal path from $(v,t)$ to any other nodes in the evolving graphs. The algorithm is shown as follows.

\begin{lstlisting}
function temporal_closeness(g::AbstractEvolvingGraph,
  start::TimeNode)
  v = start
  level = Dict(v => 0)
  i = 1
  fronter = [v]
  while length(fronter) > 0
      next = []
      for u in fronter
          for v in forward_neighbors(g, u)
              if !(v in keys(level))
                  if node_key(u) != node_key(v)
                    td = temporal_distance(start, u)
                    level[v] = i + td
                  else
                    level[v] = i - 1
                  end
                  push!(next, v)
              end
          end
      end
      fronter = next
      i += 1
  end

  total_scores = sum(values(level))
  return total_scores > 0. ?
      (length(level) - 1)/total_scores : 0.
end
\end{lstlisting}
The algorithm starts at node \texttt{start} and explores the forward neighbours and store all the
reached nodes so far in \texttt{level}. If \texttt{node\_key(u) == node\_key(v)}, we traverse the same node at different time stamps. In this case, we consider that we've already reached it at the previous time stamp.

For Figure \ref{fig:katz_eg2} the scores are
\begin{lstlisting}
TimeNode(A, 2) => 0.666667
TimeNode(A, 3) => 1.0
TimeNode(B, 1) => 0.346154
TimeNode(B, 2) => 0.0
TimeNode(B, 3) => 0.0
TimeNode(C, 1) => 0.315789
TimeNode(C, 2) => 0.5
\end{lstlisting}
and the sum of scores at different time stamps gives
\begin{lstlisting}
"A" => 1.66667
"B" => 0.346154
"C" => 0.815789
\end{lstlisting}
Note that the ranking is the same as the temporal Katz centrality case.
For Figure \ref{fig:katz_eg3}, the scores are
\begin{lstlisting}
TimeNode(A, 10)  => 0.0707071
TimeNode(A, 100) => 1.0
TimeNode(B, 1)   => 0.0597015
TimeNode(B, 10)  => 0.0
TimeNode(B, 100) => 0.0
TimeNode(C, 1)   => 0.0390625
TimeNode(C, 10)  => 0.0412371
\end{lstlisting}
and the sums are
\begin{lstlisting}
"A" => 1.07071
"B" => 0.0597015
"C" => 0.0802996
\end{lstlisting}
Similar to the temporal Katz centrality case, we see a decrease of centrality scores
for all the nodes in the evolving graph.
We note that the analysis for temporal betweenness centrality is similar to above.

@@

We see from our analysis for temporal Katz centrality and temporal betweenness centrality
that the temporal distance between nodes in an evolving graph can impact the final centrality scores.
We would like to design a temporal PageRank algorithm that takes account of between
the temporal distance and the spatial distance. The generalisation is based on two ideas:
reversed temporal walks and block adjacency matrix.

\subsubsection{Reversed Temporal Walks}

We can not travel back in time. But reversing the direction of temporal flows can help
identify the ``hub'' nodes or the source of the flows. Recall in the HITS algorithm \cite{kleinberg99},
the hub score describes the quality of a web page as a link collection of important related pages.
In reverse PageRank we compute PageRank on the graph with reversed direction, i.e., reverse the direction of each edge $(i,j)$ to $(j,i)$. Fogaras \cite{fogaras03, gleich15} shows that Reversed Page Rank scores express hub quality.

Recall a time-preserving walk is a time-ordered sequence of active nodes. We define a \emph{reversed temporal walk} to be a reversed time-ordered sequence of active nodes
$\langle (v_1, t_1), (v_2,t_2), \ldots, (v_m, t_m)\rangle$, where
$t_1 \ge t_2 \ge \cdots \ge t_m$.

@@

Aggregating edges at each time stamp to form
a simple static graph can provide misleading information about the structure of the network.
In \cite{chen16}, we derive a block adjacency matrix representation of an evolving graph and prove it is the correct representation of an evolving graph.
Let $E'$ is the set of causal edges and $\hat E$ is the set of static edges then
an evolving graph can be represented as
$$
\bm M_n =
\begin{pmatrix}
A^{[t_1]} & M^{[t_1, t_2]} & \ldots & M^{[t_1, t_n]} \\
0         & A^{[t_2]} & \ldots & M^{[t_2, t_n]} \\
          & \ldots    &        &     \\
0         & 0         & \ldots & A^{[t_n]}
\end{pmatrix},
$$
where $M^{[t_i, t_j]}$ is the matrix whose rows are labeled by $V^{[t_i]}$ and columns are labeled by $V^{[t_j]}$, and whose entries are
$$
  M_{uv}^{[t_i, t_j]} =
  \begin{cases}
    1 & \mbox{if} (u, v) \in E' \\
    0 & \mbox{otherwise}.
  \end{cases}
$$
The adjacency matrix blocks $A^{[t]}$ encode the static edge set $\hat E$, whereas the off-diagonal blocks $M^{[t_i, t_j]}$ encode the causal edge set $E'$. To take account of the temporal distance between nodes, we could let
$$
  M_{uv}^{[t_i, t_j]} =
  \begin{cases}
    \frac{1}{|t_j - t_i|} & \mbox{if} (u, v) \in E' \\
    0 & \mbox{otherwise}.
  \end{cases}
$$

For Figure \ref{fig:eg_shortest_walk}, we have

\begin{align*}
V & = \{(1,\!1998), (2,\!1998), (1,\!1999), (3,\!1999), (2,\!2018), (3,\!2018)\},\\
\tilde E &= \{((1,\!1998), (2,\!1998)), ((1,\!1999), (3,\!1999)), ((2,\!2018), (3,\!2018))\},\\
E'  &= \{((1,\!1998), (1,\!1999)), ((2,\!1999), (2,\!2018)), ((3,\!1999), (3,\!2018))\}.
\end{align*}

@@

\[
\bm M_3 = \begin{pmatrix}
0 & 1 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1/9 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1/9 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 0
\end{pmatrix}
\]

@@

Notice that the lower triangular blocks of matrix $\bm M_n$ are zeros.
This is because all the causal edges are time preserving. We consider the transpose of matrix $\bm M_n$, which reverse the direction of static edges and causal edges.
Let $b \in [0,1]$ be a balance scalar. Then the block matrix
$$
b \bm M_n + (1-b) \bm M_n^T
$$
takes account of both the temporal walks and reversed temporal walks.
We now derive the block Google matrix as

\begin{lstlisting}
function block_google_matrix(g; alpha = 0.85, balance = 0.75)
  M = full(block_adjacency_matrix(g))
  M = balance * M + (1-balance) * M'
  N, N = size(M)
  p = ones(N)/N
  dangling_weights = p
  dangling_nodes = find(x->x==0, sum(M,2))
  for node in dangling_nodes
      M[node,:] = dangling_weights
  end
  M ./= sum(M,2)
  return alpha * M .+ (1- alpha) * p
end
\end{lstlisting}

The function \texttt{block\_adjacency\_matrix} convert an evolving graph $g$ to a block adjacency matrix $\bm M$. Every row sum of the block Google matrix is equal to one. Each element of the dominant eigenvector of the block Google matrix
represents the PageRank rating of a node.

For Figure \ref{fig:katz_eg2}, we set $alpha$ be $0.85$ and $balance$ be $1.0$.
Then the PageRank scores are
\begin{lstlisting}
 TimeNode(A, 2) => 0.289123
 TimeNode(A, 3) => 0.128054
 TimeNode(B, 1) => 0.182477
 TimeNode(B, 2) => 0.354335
 TimeNode(B, 3) => 0.480941
 TimeNode(C, 1) => 0.128054
 TimeNode(C, 2) => 0.250931
\end{lstlisting}
and the sum of node scores at all three time stamps are
\begin{lstlisting}
 "A" => 0.417177
 "B" => 1.01775
 "C" => 0.378986
\end{lstlisting}
We see node $B$ is the most important node in an evolving graph.
On the other extreme, if we set $balance$ to be $0.0$, we get the overall scores
\begin{lstlisting}
 "A" => 0.678502
 "B" => 0.647353
 "C" => 0.527927
\end{lstlisting}
Thus $A$ is the most important node in this case.

One potential problem with temporal PageRank is that the computational cost is expensive for large evolving graphs with many time stamps. The dimension of the block matrix
is $|V| \times |T|$, where $|V|$ represents the number of nodes and $|T|$ represents the number of time stamps.
\section{Experiments}
\label{sec:experiments}

We test the influence of time on evolving graph centralities.
Justify use \texttt{IntAdjacencyList}.
